---
title: "The global implications of taxing land in agricultural production: Land use, agricultural production, and consumer well-being"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Eric C. Davis, Maros Ivanic, and Brent Sohngen"
output: 
  bookdown::pdf_document2:
    keep_tex: TRUE
toc: FALSE
bibliography: references.bib
header-includes:
    - \usepackage{setspace}\doublespacing
abstract: "The projected global growth in population and incomes is expected to create pressure to convert forestland into farmland, with an estimated 144.2 million hectares of forestland estimated to be converted to agricultural land by 2033. At the same time, increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink, but taking such actions would likely result in less agricultural output and higher food prices. Using the GTAP and GTM models, we model how different tax rates on the value of output from land used in agricultural production would impact deforestation. We find that, with a tax rate of zero percent on agricultural production, 144.2 million hectares (3.6%) of global forests would be converted to agricultural lands by 2033. With a 70 percent tax, however, the net deforestation rate would be brought to zero while the well-being of global consumers would be reduced by $92.1 billion in 2033, which translates to a global average annual cost of $10.6 per person."
---

```{r setup, include=FALSE, cache = TRUE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\vspace{1in}

**The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S. Government determination or policy.**

**This research was supported in part by the U.S. Department of Agriculture, Economic Research Service.**

\newpage

# Introduction

In 2020, there were 4.1 billion hectares of forestland globally covering roughly 31.4% of the land [@fao2023]. These forests play an important role in maintaining and growing the land-based carbon sink by sequestering about 30% of GHG emissions currently [@roe2019]. This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021]. With increasingly negative climate impacts, there is increased attention being placed on maintaining these carbon sinks, but this requires, at minimum, avoiding further deforestation. However, from 2015 to 2020, roughly 10 million hectares of forestland was converted annually to other land uses [@fao2023]. There are many reasons for this. As nations develop and disposable incomes rise, for example, consumers often increase meat consumption, which requires a greater quantity of agricultural land than is needed for vegetable-based diets [@elferink2007]. Population growth is also a driver. Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the world's population will grow by 8.8 percent, a gain of 707.1 million people [@un2024], and where population growth is especially high, which is the case at present throughout Africa (2023 to 2033 population growth: 25.0%: 360.6 million), there is likely to be greater conversion of forestland into farmland. This puts into question whether avoiding further deforestation is attainable.

Moreover, if efforts to end deforestation are to be made, it is important to understand where to target them, as forests do not provide equal carbon benefits in every region. Coniferous forests in boreal zones are known to provide the smallest net benefits, as gains in carbon sequestration are mitigated by reductions in albedo in winter [@mykleby2017; @li2015]. These forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021]. Conversely, tropical forests have been linked with a strong positive results, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015]. 

As land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale. Several notable modeling attempts have been made in this direction. @steinbuks2012 developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects [@steinbuks2012; @hertel2017]. Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes. This prompted the creation of the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products [@baldos2012]. The GTAP Agro-Ecological Zone (GTAP-AEZ) model further strengthened the GTAP model's ability to analyze land use changes, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates. 

More recent modeling attempts of land-use change have begun to move away from comparative static analysis. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by @stevenson2013. This model has been utilized to evaluate the impact of crop intensification on land use. It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014]. In addition, in the standard GTAP model, it is difficult to model a sector individually. There are 68 market sectors covered in the latest version of the GTAP database. The GTAP model's general equilibrium nature means that all these markets are cleared with market-clearing prices. As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model. 

This project gets around these concerns and assesses the impact of various tax scenarios on land-use by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001]. The latest versions of both models were used. For the GTAP model, this meant using data which has 2017 as its base year. This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) [@CEPII2024] to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world. Similar steps were taken to update and recalibrate the GTM model. As most forest inventory data are not updated annually, the GTM model updates data for regions with inventory information approximately every five years, and it updates data for other regions using the Global Forest Resource Assessment [@fao2020]. 


# Connecting the GTAP and GTM models

The GTAP and GTM models are both widely used for policy analysis. The GTAP model is frequently used to describe the implications of various policy instruments on national and global markets and their prices, output, and trade. The GTM model is often used to estimate the impacts of changes in population, GDP, and land rents on global forest stocks and carbon sequestration. While both models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, as both consider several important variables exogenous. By combining the two models, we expand each model's descriptive power and thus produce a better depiction of the global economy, which should allow for a more precise assessment of the impact of policy actions. 

Specifically, we use the GTAP model to detail the impact on land rents based on the amount of available agricultural land. This benefits the GTM model, which uses returns as an exogenous input. We then use the GTM model to determine the change in agricultural land quantity based on the returns to land. This benefits the GTAP model, which takes the change in the amount of land available for crop and livestock production as an exogenous input. Using these links, we are able to solve both models concurrently, and outputs are passed back and forth between the models until we achieve a dual-state equilibrium. In the GTAP model, we solve for the variable `qo` for land, and in the GTM model, we solve for the parameter `RENTA`. These are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model mirrors the associated change in agricultural land (calculated as the non-forested area from the total available land) from the change in rents in GTM. 

On the software level, we establish the connection between the models using R software [@R2021]. This is required as the two models were written using using different languages. Moreover, GTAP, which was written in GEMPACK, and GTM, which was written in GAMS, were developed long before software integration became important or even possible, and so we use an R script, which is capable of handling other software, including generating required inputs, executing them, and reading in any outputs, to create the user functions that execute both models, collect their outputs, and automatically create any required input files.  

<!-- Once R was able to control the interactions between both models, we were then able to formulate the combined solution as a fairly simple math program that is solved in R: -->

<!-- \begin{equation} -->
<!-- \mbox{GTAP}:V_1\to V_2 -->
<!-- \end{equation} -->

<!-- \begin{equation} -->
<!-- \mbox{GTM}:V_2\to V_1 -->
<!-- \end{equation} -->

<!-- \begin{equation} -->
<!-- s.t. V_2 = V_1 -->
<!-- \end{equation} -->

<!-- where $V_1$ is a vector of all the variables in both systems, i.e., all variables in GTAP and all variables in GTM. Of import, $V_1$ includes the land change variables, which are exogenous to the GTAP model. $V_1$ is then used by the GTAP model (represented by function \mbox{GTAP}) and its outputs are passed to GTM as a vector of variables $V_2$. This output vector contains the change in rents, which is treated as exogenous in the GTM model. GTM then processes the $V_2$ variables as inputs and passes its solution back to GTAP as $V_1$. To solve the system simultaneously, we require that $V_1$ and $V_2$ be equal. We focus on the relative returns of land in agriculture and forestry as they greatly impact land use changes. Thus, in our scenarios, we examine the impact of different tax rates on the use of land in the production of agricultural output. With several iterations, the system converges to the change in land and land rent that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model. -->


# Changes under the zero tax scenario

Our first scenario examines the likely changes over the decade under study, 2023--2033. To build our scenario, we use data from CEPII [@CEPII2024] and focus on two of their variables that we determined to be key: population and GDP. In Table \@ref(tab:baseline-growth), we show the projected population and GDP growth per region. GDP increases for all regions of the world with the largest percentage changes occurring in South Asia (+308.6%), China (+189.8%), and Sub-Saharan Africa (+172.9%). Population also is projected to grow in most regions with the largest changes happening in Sub-Saharan Africa (+55.7%). Japan, Russia, and East Asia are projected to see their populations decline by 10.4%, 4.7%, and 2.6%, respectively. As the CEPII projections do not contain any information on land use, we also include the projections on forested land change included in the GTM. Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we then calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033 while assuming a uniform zero percent tax on the use of land in the production of agricultural output. Results suggest that there will be a significant increase in agricultural land in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%). Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to population decline. 

```{r echo = FALSE}
regLabels = c(
  "US" = 'United States',
  "CHINA" = 'China',
  "BRAZIL" = 'Brazil',
  "CANADA" = 'Canada',
  "RUSSIA" = 'Russia',
  "EU ANNEX I" = 'West/Central Europe',
  "EU NON ANNEX I" = 'EU II',
  "SOUTH ASIA" = 'South Asia',
  "CENT AMER." = 'Central America',
  "RSAM" = 'Other Latin America',
  "SSAF" = 'Sub-Saharan Africa',
  "SE ASIA" = 'SE Asia',
  "OCEANIA" = 'Oceania',
  "JAPAN" = 'Japan',
  "AFME" = 'North Africa/Middle East',
  "E ASIA" = 'East Asia'
)

gtapCountries = HARr::read_har('asset/gtap/flexagg/gtm/gsdgset.har')$h1


gtmCountries = c(
  'US',
  'CHINA',
  'BRAZIL',
  'CANADA',
  'RUSSIA',
  'EU ANNEX I',
  'EU NON ANNEX I',
  'SOUTH ASIA',
  'CENT AMER.',
  'RSAM',
  'SSAF',
  'SE ASIA',
  'OCEANIA',
  'JAPAN',
  'AFME',
  'E ASIA'
)

gGTAP = gtapCountries
names(gGTAP)=c("OCEANIA","CHINA","JAPAN","E ASIA", "SE ASIA", "SOUTH ASIA", "CANADA","US",  "CENT AMER.","BRAZIL", "RSAM",  "EU ANNEX I", 
               "EU NON ANNEX I",  "RUSSIA", "SSAF", "AFME", "ROW")

gGTM = names(gGTAP)
names(gGTM)= gGTAP


gGTM['ROW']='ROW'
regLabels['ROW']='ROW'

baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')

out = cbind('Real GDP (\\%)' = baselineGTAP$qgdp[,'TOTAL'],'Population (\\%)' = baselineGTAP$pop[,'TOTAL'],'Agricultural land (\\%)' = baselineGTAP$qo['land',,'TOTAL'])

rownames(out) = regLabels[gGTM[rownames(out)]]

rownames(out)[length(rownames(out))] = 'ROW'


kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'GDP, population, and agricultural land growth under the baseline scenario (Source: CEPII and GTM)', digits = 1, escape = FALSE, label = 'baseline-growth')
  ), 1, border_left = TRUE),
  4,
  border_right = TRUE
)
```

Focusing on the absolute area of the forestland, as generated by the GTM model, we estimate that under this scenario between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)). Southeast Asia, Brazil, and the rest of South America also see forestland shrink, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively. Conversely, the United States is estimated to add 1.3 million hectares (0.5%) to its forestland.

```{r echo = FALSE}


baseline = readRDS('history_baseline.rds')

out = baseline$gtmResult$forests[1:2,]

rownames(out) = c('Forestland in 2023','Forestland in 2033')

out = rbind(out, `$\\Delta$ (Mha)` = out[2,]-out[1,],`$\\Delta$ (\\%)` = 100*(out[2,]/out[1,]-1))
colnames(out) = regLabels[colnames(out)]

out = cbind(out, Total = rowSums(out))

out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100


kableExtra::column_spec(
  kableExtra::column_spec(
    kableExtra::kable_styling(
      kableExtra::add_footnote(
        knitr::kable(
          t(out),
          escape = FALSE,
          #format = 'latex',
          digits = 1,
          label = 'baseline-forest-change',
          caption = 'Forestland by region in 2023 and estimated change in 2033 relative to 2033 baseline scenario'
        ),
        label = c(
          'a' = 'Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.',
          'b' = 'Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.',
          'c' = 'Rest of South America includes Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.',
          d = 'Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.',
          e = 'Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.'
        ),
        threeparttable  = TRUE
      )
    ),
    1,
    border_left = TRUE
  ),
  5,
  border_right = TRUE
)
```

# Modeling the impact of higher tax scenarios

Our second scenario examines the impact of a tax of roughly 70 percent on the use of agricultural land. Solving the two models together shows that the economic impact would not be limited to the cost of the tax, it would also impact food production and its prices. For example, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed. For the United States, there would be 2.1 percent less agricultural land (Table 3). Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-7.9 percent), Brazil (-8.6 percent), and Russia (-9.3 percent). This would be because the global net deforestation rate would be brought to roughly zero, meaning that the 144.2 million hectares that would have been converted to agricultural land by 2033 in the zero tax scenario would be preserved as forestland.

This would have impacts on the quantity of agricultural goods that could be produced. For example, the 15.0 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in reductions in agricultural production of 5.0 percent and 1.7 percent, respectively. Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.6 percent and 1.0 percent, respectively. This is the result of the sharp output declines in other regions prompting increased intensity of production in the United States and Brazil. When we examine how the policy impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4). 



```{r echo = FALSE}

baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')

out = cbind(
  'Real GDP (\\%)' = ((1 + worldGTAP$qgdp[, 'TOTAL'] / 100) / (1 + baselineGTAP$qgdp[, 'TOTAL'] /
                                                                 100) - 1) * 100,
  'Agr. land (\\%)' = ((1 + worldGTAP$qo["land",, 'TOTAL'] / 100) / (1 + baselineGTAP$qo["land",, 'TOTAL'] /
                                                                 100) - 1) * 100,
  'Agr. prices (\\%)' = ((1 + worldGTAP$pm["ag",, 'TOTAL'] / 100) / (1 + baselineGTAP$pm["ag",, 'TOTAL'] /
                                                                 100) - 1) * 100,
  'Agr. output (\\%)' = ((1 + worldGTAP$qo["ag",, 'TOTAL'] / 100) / (1 + baselineGTAP$qo["ag",, 'TOTAL'] /
                                                                 100) - 1) * 100
)
rownames(out) = regLabels[gGTM[rownames(out)]]

rownames(out)[length(rownames(out))] = 'ROW'


kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Changes in key variables in 2033 under a uniform land tax of 70 and relative to 2033 baseline scenario \\%', digits = 1, escape = FALSE, label = 'baseline-scenario1')
  ), 1, border_left = TRUE),
  5,
  border_right = TRUE
)

```


```{r echo = FALSE}


world = readRDS('history_world.rds')

out = world[[8]]$gtmResult$forests[1:2,]

rownames(out) = c('Forestland in 2023','Forestland in 2033')

out = rbind(out, `$\\Delta$ (Mha)` = out[2,]-out[1,],`$\\Delta$ (\\%)` = 100*(out[2,]/out[1,]-1))
colnames(out) = regLabels[colnames(out)]

out = cbind(out, Total = rowSums(out))

out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100


kableExtra::column_spec(
  kableExtra::column_spec(
    kableExtra::kable_styling(
        knitr::kable(
          t(out),
          escape = FALSE,
          #format = 'latex',
          digits = 1,
          caption = 'Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place relative to 2033 baseline scenario (Source: Authors\' calculations)'
        )
    ),
    1,
    border_left = TRUE
  ),
  5,
  border_right = TRUE
)


```

A visual depiction of how this tax would impact global trade is presented in Figure 1. The United States would be expected to see its agricultural exports increase significantly. Conversely, the regions where agricultural exports would be expected to drop the most are East Asia, South Asia, Southeast Asia, and Central America, which would see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.

```{r echo= FALSE, fig.cap='Changes in ag trade volumes (red arrows denote reductions, green arrows denote increases)'}
baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')
baselineGTAPC = HARr::read_har('asset/GTAP/run_baseline.slc', TRUE)


changesInAgTrade = ((1+worldGTAP$qxs/100)/(1+baselineGTAP$qxs/100)-1)['ag',,,'TOTAL'] * baselineGTAPC$viws['ag',,]

colnames(changesInAgTrade) = rownames(changesInAgTrade) = regLabels[gGTM[c("ROW", "china", "japan", "ROW", "seasia", "ROW", "ROW", "usa", "ROW", "brazil", "samerica", "eui", "ROW", 
"ROW", "ssa", "ROW", "ROW")]]



changesInAgTrade = matsbyname::aggregate_byname(changesInAgTrade)

diag(changesInAgTrade)=0


#round(changesInAgTrade,0)

circlize::chordDiagram(changesInAgTrade, directional = TRUE, col = ifelse(changesInAgTrade>0,'green','red'),direction.type = c("arrows"), link.arr.type = 'big.arrow')

```

Overall, the 70 percent tax, which results in zero net deforestation globally, is likely to have a negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
$977.1 million and $166.9 million, respectively (Table 5). Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes. The United States, along with 10 other regions, are estimated to see well-being drop by more than $1 billion. The most negatively impacted regions would be South Asia (-$28.0 billion), China (-$14.5 billion), and Southeast Asia (-$13.6 billion). In total, this tax-driven policy is estimated to decrease global consumer well-being in 2033 by $92.1 billion relative to the scenario where no tax on agricultural land was in place. 

```{r echo = FALSE}

baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')


out = cbind(
  `Baseline (B\\$)` = baselineGTAP$ev[, 'TOTAL'] / 1000,
  `Scenario (B\\$)` = worldGTAP$ev[, 'TOTAL'] / 1000,
  `$\\Delta$ (B\\$)` = (worldGTAP$ev[, 'TOTAL'] - baselineGTAP$ev[, 'TOTAL']) /
    1000
)

rownames(out) = regLabels[gGTM[rownames(out)]]

#rownames(out)[length(rownames(out))] = 'ROW'

out = rbind(out, 'Total'=colSums(out))

kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in billions of U.S. dollars, with deforestation mitigation policy in place relative to 2033 baseline scenario', digits = 1, escape = FALSE, label = 'EV',format.args = list(big.mark = ","))
  ), 1, border_left = TRUE),
  4,
  border_right = TRUE
)

```

Much of this loss in well-being may be the result of higher prices. Japan is the only nation that is estimated to see any price impacts that are beneficial for consumers, with the price of agricultural goods decreasing 0.3 percent (Table \@ref(tab:pm)). All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent. 

```{r echo= FALSE}



baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')

out = data.frame('Ag prices'=(1+worldGTAP$pm['ag',,'TOTAL']/100)/(1+baselineGTAP$pm['ag',,'TOTAL']/100)-1, check.names = FALSE) *100

rownames(out) = regLabels[gGTM[rownames(out)]]
#out = cbind(out, Total = rowSums(out))
#out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100

kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, with deforestation mitigation policy in place, relative to 2033 baseline scenario', digits = 1, escape = FALSE, label = 'pm')
  ), 1, border_left = TRUE),
  2,
  border_right = TRUE
)

```

# Discussion

This analysis shows that tax rates on agricultural land can have a significant impact on land use. At a global average tax rate of zero percent of the value of land used in agricultural production, an estimated 144.2 million hectares of forestland would be converted to agricultural land by 2033. With an estimated global tax rate of 70.3 percent, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland. The economic price of this would likely be transferred to consumers through lower agricultural output and higher market prices. Overall, this land tax is estimated to reduce consumer well-being by $92.1 billion in 2033, which equates to $10.60 per person annually. 

This work also shows the utility of joining two major economic models. GTAP has long been valued for its ability to provide robust information on global trade flows and prices, and GTM has proven its utility in understanding the impact of forestry decisions on the carbon sink and land rents. By combining these models through the use of R software and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized. This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their inter-dependencies.


# Conclusions

According to CEPII’s projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied by a moderate growth in population. Another set of projections by GTM suggests that this growth would also mean significant reductions in the world’s forests because they would be converted to agricultural land, with the model predicting a loss of over 144 million hectares of forest by 2033. By combining two well-known global models—the GTAP model that describes the world markets and trade, and the GTM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas—we were able to estimate the impact of a tax on agricultural land on land use change. We find that a tax of about 70 percent would likely prevent net global deforestation in the coming ten years through reducing the returns to land when used in agriculture. We also observe that the reduction in available agricultural land relative to the baseline growth would likely result in reduced agricultural output and higher prices, which might have food security implications. Finally, we note that such a uniform global tax that would reduce the returns to land to the point of preventing global deforestation appears to produce highly differentiated impacts around the world’s regions. Some regions, such as China, Russia, Canada, and the United States are estimated to see their forested areas increase as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil, and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area. 





\newpage

# References








<!-- Unused -->

<!-- 1a) On aggregate zero percent net deforestation Globally -->

<!-- 1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023]. To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa". -->

<!-- 2) Partial attainment of AFR100 pledges as part of Bonn Challenge: This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating. Using this measure, there would only be a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023]. To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa". -->

<!-- 3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023]. To see the list of countries and the hectares pledged, please view the Latin America tab of the Excel document "Bonn Challenge-Other Regions". -->

<!-- 4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023]. To see the list of countries and the hectares pledged, please view the Tropical Forest tab of the Excel document "Bonn Challenge-Other Regions". Nations that lie in both tropical and sub-tropical zones and taken to be tropical. -->

<!-- 5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020]. This will be modeled as a uniform 3% increase in forested area across all nations. -->

<!-- 6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020]. This will be modeled as a uniform 15% increase in forested area across all nations. -->

<!-- [@defries2010] found a strong correlation between forest loss and exports of agricultural products. On the other hand, [@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources. -->

